### 2017-03-22 doc/manuscript.Rmd
Added abstract

---
bibliography: 
    - /home/luke/Documents/bibliography/bib/bib-db/tagDiabetes.bib
    - /home/luke/Documents/bibliography/bib/bib-db/myrefs.bib
    - /home/luke/Documents/bibliography/bib/bib-db/methods.bib
csl: /home/luke/Documents/bibliography/styles/dependent/diabetes-care.csl
output: fost::diabetologia_manuscript
---

@@

>>>>

---
bibliography: 
    - /home/luke/Documents/bibliography/bib/bib-db/tagDiabetes.bib
    - /home/luke/Documents/bibliography/bib/bib-db/myrefs.bib
    - /home/luke/Documents/bibliography/bib/bib-db/methods.bib
csl: /home/luke/Documents/bibliography/styles/dependent/diabetes-care.csl
output: fost::diabetes_care_manuscript
---

@@

Word count: ~281 / 250

**Objective**: 
There is limited data on the role of the specific triacylglyceride fatty acid
(TAGFA) composition on diabetes pathogenesis. Our aim was to examine the 
longitudinal association of TAGFA on insulin sensitivity (IS) and beta-cell
function.

**Research Design and Methods**:
We used data from the Prospective Metabolism and Islet Cell Evaluation (PROMISE)
cohort of adults (n=477) at-risk for diabetes with over 6-years of followup (3
timepoints). Glucose and insulin from an OGTT were used to calculate the Matsuda
index (ISI), HOMA2-%S, the Insulinogenic Index over HOMA-IR (IGI/IR), and the
Insulin Secretion-Sensitivity Index-2 (ISSI-2). Gas chromatography quantified
TAGFA. Generalized estimating equations (GEE) adjusted for confounders and
partial least squares (PLS) were used for the analysis.

**Results**: 
The outcome variables declined by 14% to 27% over the 6-years. Most TAGFA
modeled as a concentration had strong negative associations with IS in adjusted
GEE models; when as a mol%, four TAGFA (14:0, 16:0, 14:1n-7, 16:1n-7) had strong
negative associations while most others (e.g. 18:1n-7, 18:1n-9, 20:2n-6,
20:5n-3) had strong positive associations. Few associations were seen for
beta-cell function, except for 16:0 (negative) and 18:1n-7 (positive). The PLS
analysis identified that the four 'harmful' TAGFA (14 and 16 carbon TAGFA)
clustered together and strongly predicted lower IS.

**Conclusions**:
We found that a higher proportions of a cluster of four TAGFA (14:0, 14:1-7,
16:0, 16:1n-7) strongly predicted lower IS. Previous research suggests these
four FA are heavily derived from the de novo lipogenesis (DNL) of higher intakes
of simple or refined carbohydrates, especially 16:1n-7. Our results suggest that
greater simple or refined carbohydrate intake may impact risk for diabetes
through greater DNL and higher FA associated with lipotoxicity (e.g. 16:0),
leading to hypertriglyceridemia and lower IS.
